Global Life Expectancy: Unraveling Health and Economic Determinants (2015-2020)

### 2024-12-04 paper/paper.qmd
commint

According to the @tbl-sum_sta, we could tell that

>>>>

According to @fig-summary, we could tell that the mean life expectancy over the 6 years is 72.2, with a relatively low standard deviation of 7.8. This suggests that, on average, life expectancy across the dataset is fairly consistent, with only moderate variation from the mean. However, the maximum life expectancy of 87.4 stands out significantly, indicating that some countries or regions are achieving substantially higher life expectancy.With the mean at 72.2 and the median at 73.1, it’s apparent that the data is fairly symmetric, with the median being slightly higher than the mean, suggesting a mild left skew. This shift, combined with the large maximum value, hints at the presence of outliers—countries with exceptionally high life expectancy that pull the maximum up. The variance of 60.1 is relatively large, further supporting the idea that the data spans a wide range, with some observations considerably deviating from the central tendency.

Given this, the high variance can be attributed to the diversity in life expectancy values across different regions, particularly when countries at the higher end of the spectrum (e.g., those with life expectancy around 87 years) are compared with those at the lower end. This variability suggests that while most countries have life expectancy near 72 years, the outliers significantly influence the spread of the data, leading to a higher variance than what would be expected in a more homogenous dataset. Therefore, while the standard deviation is relatively small, the data is far from uniform, and a deeper analysis of the outliers could provide further insights into the factors contributing to the high life expectancy in certain regions.

@tbl-sum_sta summarizes global life expectancy across various categories such as gender, geographic regions, and income levels. We could tell that the global life expectancy is 72.13 years, providing a benchmark against which other categories can be compared. Women have a significantly higher life expectancy (74.62 years) than men (69.71 years), which reflects a common global trend due to biological and social factors. Region variation does happens to influence the life expectncy. Europe has the highest life expectancy (78.50 years), which is likely due to better healthcare and living conditions, with having the most amount of high income countries. Africa has the lowest (64.17 years), potentially reflecting challenges like limited healthcare infrastructure, poverty, and infectious diseases. Income level, as one of the most essential factor, significantly affect the life expectancy. High-income countries have the highest life expectancy (79.38 years), illustrating the correlation between economic wealth and longevity. High income level are believed to enable better living standard compared to other income groups. Upper middle income countries has an average of 73.38 years, where we could observe that the difference is significant lower than high income group. Lower middle income group has the life expectancy drop to 68.40. While these countries usually face challenges such as infrastructure and limit access to health resources, economic growth may need to be taken seriously be policy makers. Lower-income countries have a significantly lower life expectancy of 62.35 years, which can be attributed to limited access to healthcare, education, and basic services. @tbl-sum_sta highlights global inequalities and emphasizes the importance of improving living standards, especially in lower-income and underdeveloped regions.
